Imputation of Missing Values in the Melbourne CBD Pedestrian Dataset

### 2017-09-03 thesis/missing_impute.Rmd
southbank model

---
title: "Imputation of Missing Values in the Melbourne CBD Pedestrian Dataset"
output:
  html_document:
    theme: united
  html_notebook: default
  pdf_document: default
---
```{r, include=FALSE}
library(plyr)
library(tidyverse)
library(lubridate)
library(pander)
library(gridExtra)
library(fortify)
set.seed(12345)
```

@@

@@

### Standardise Counts

@@

@@

flin_swan <- lm(swan ~ liz + underpass, data = flin_swan_df) 
summary(flin_swan)
flin_swan_df$swan_fit <- predict.lm(flin_swan, newdata = flin_swan_df)
flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
plot(flin_swan)
ggplot(flin_swan_df) + 
  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')

@@

flin_swan <- lm(swan ~ daytype*liz*timefact, data = flin_swan_df) 
summary(flin_swan)
flin_swan_df$swan_fit <- predict.lm(flin_swan, newdata = flin_swan_df)
flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
plot(flin_swan)
ggplot(flin_swan_df) + 
  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')

@@

```{r}
flin_swan_2 <- lm(swan ~ daytype*timefact*liz + daytype*timefact*underpass, data = flin_swan_df) 
summary(flin_swan_2)
flin_swan_df$swan_fit <- predict.lm(flin_swan_2, newdata = flin_swan_df)
flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
plot(flin_swan_2)

@@

flin_swan_2 <- lm(swan ~ timefact*liz + timefact*underpass, data = flin_swan_df_train) 
summary(flin_swan_2)
flin_swan_df$swan_fit <- predict.lm(flin_swan_2, newdata = flin_swan_df)
flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
plot(flin_swan_2)

@@

```{r}
flin_swan_df[8763:8764, ]
```

High leverage points here are the New Year's Day celebrations.

>>>>

---
title: "Imputation of Missing Values in the Melbourne CBD Pedestrian Dataset"
output:
  html_document:
    theme: united
  html_notebook: default
  pdf_document: default
---
```{r, include=FALSE}
library(plyr)
library(tidyverse)
library(lubridate)
library(pander)
library(gridExtra)
library(ggfortify)
set.seed(12345)
```

@@

for (i in 1:43) {
  ped_list[[i]]$glm_fitted <- predict.glm(ped_models[[i]], ped_list[[i]], type = "response")
}

@@

### Standardising Counts

@@

#### Flinders St. - Swanston St. (West)

@@

flin_swan <- lm(swan ~ liz + underpass, data = flin_swan_df) 
summary(flin_swan)
flin_swan_df$swan_fit <- predict.lm(flin_swan, newdata = flin_swan_df)
flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
autoplot(flin_swan)
ggplot(flin_swan_df) + 
  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')

@@

flin_swan <- lm(swan ~ daytype*liz*timefact, data = flin_swan_df) 
summary(flin_swan)
flin_swan_df$swan_fit <- predict.lm(flin_swan, newdata = flin_swan_df)
flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
autoplot(flin_swan)
ggplot(flin_swan_df) + 
  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')

@@

```{r}
flin_swan_2 <- lm(swan ~ daytype*timefact*liz + daytype*timefact*underpass, data = flin_swan_df) 
summary(flin_swan_2)
flin_swan_df$swan_fit <- predict.lm(flin_swan_2, newdata = flin_swan_df)
flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
autoplot(flin_swan_2)
ggplot(flin_swan_df) + 
  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')

@@

flin_swan_2 <- lm(swan ~ timefact + timefact:liz + timefact:underpass, data = flin_swan_df_train) 
summary(flin_swan_2)
flin_swan_df$swan_fit <- predict.lm(flin_swan_2, newdata = flin_swan_df)
flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
autoplot(flin_swan_2)

@@

```{r}
flin_swan_df[c(8763:8764, 1248), ]
```

High leverage points here are the New Year's Day celebrations.  
Comparing this fit to that Quasipoisson GLM:
```{r}
flin_swan_var <- ped_list$`Flinders St-Swanston St (West)` %>%
  .$Hourly_Counts %>% var(na.rm = TRUE)
flin_swan_mu <- ped_list$`Flinders St-Swanston St (West)` %>%
  .$Hourly_Counts %>% mean(na.rm = TRUE)
flin_swan_df$glmfits <- ped_list$`Flinders St-Swanston St (West)` %>% .$glm_fitted
flin_swan_df$fits <- flin_swan_df$swan_fit*sqrt(flin_swan_var) + flin_swan_mu
flin_swan_df$swan_unscaled <- flin_swan_df$swan*sqrt(flin_swan_var) + flin_swan_mu

ggplot(flin_swan_df) + 
  geom_path(aes(x = time, y = swan_unscaled, group = date), alpha = 0.05) + 
  geom_path(aes(x = time, y = glmfits, group = date), alpha = 0.02, colour = 'blue') + 
  geom_path(aes(x = time, y = fits, group = date), alpha = 0.02, colour = 'red')

#### Southbank
```{r}
ped_df[ped_df$Sensor_ID == "Southbank", ]$scaled_count -> sbank
ped_df[ped_df$Sensor_ID == "Waterfront City", ]$scaled_count -> wfront
ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$scaled_count -> wbridge
ped_df[ped_df$Sensor_ID == "Vic Arts Centre", ]$scaled_count -> vicarts

sbank_df <- data.frame(sbank_y = sbank, wfront, wbridge, vicarts,
                       date = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Date,
                       time = as.integer(ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Time),
                       timefact = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Time,
                       month = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Month,
                       daytype = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$DayType,
                       dtm = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Date_Time,
                       hday = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$IsHDay)

qplot(x = dtm, y = sbank, data = sbank_df)
```

```{r}
ggplot(sbank_df) + geom_point(aes(x = sbank, y = wbridge))

sbank_fit <- lm(sbank_y ~ wfront + wbridge + vicarts, data = sbank_df)
summary(sbank_fit)
autoplot(sbank_fit)
sbank_df$sbank_fit <- predict(sbank_fit, sbank_df)
ggplot(sbank_df) + geom_path(aes(x = time, y = sbank_y, group = date), alpha = 0.06) + 
                   geom_path(aes(x = time, y = sbank_fit, group = date), alpha = 0.02, colour = 'red')
```

```{r}
sbank_fit <- lm(sbank ~ timefact:wbridge + timefact:vicarts + timefact:wfront, data = sbank_df)
summary(sbank_fit)
autoplot(sbank_fit)
sbank_df$sbank_fit <- predict(sbank_fit, sbank_df)
ggplot(sbank_df) + geom_path(aes(x = time, y = sbank_y, group = date), alpha = 0.06) + 
                   geom_path(aes(x = time, y = sbank_fit, group = date), alpha = 0.02, colour = 'red')
```

```{r}
sbank_df[26302:26304, ]
```

We also have a huge potential outlier:
```{r}
sbank_df %>% arrange(desc(fits)) %>% head(5) %>%
    select(dtm, daytype, sbank_y, wfront, vicarts,
            wbridge, sbank_fit, fits, glmfits) %>%
    pander()
```

Comparing to the Quasipoisson GLM:
```{r}
sbank_var <- ped_list$`Southbank` %>%
  .$Hourly_Counts %>% var(na.rm = TRUE)
sbank_mu <- ped_list$`Southbank` %>%
  .$Hourly_Counts %>% mean(na.rm = TRUE)
sbank_df$glmfits <- ped_list$`Southbank` %>% .$glm_fitted
sbank_df$fits <- sbank_df$sbank_fit*sqrt(sbank_var) + sbank_mu
sbank_df$sbank_unscaled <- sbank_df$sbank_y * sqrt(sbank_var) + sbank_mu

ggplot(sbank_df[sbank_df$date != "2014-04-13", ]) + 
  geom_path(aes(x = time, y = sbank_unscaled, group = date), alpha = 0.15) + 
  geom_path(aes(x = time, y = glmfits, group = date), alpha = 0.007, colour = 'blue') +
  geom_path(aes(x = time, y = fits, group = date), alpha = 0.0088, colour = 'red')
```
